Predicting the manner in which an exercise was performed.
---------------------------------------------------------

```{r}
library(doMC)
library(caret)
library(randomForest)

setwd("/home/anon/")
```
Reading in the data
```{r}
training = read.csv("education/practical machine learning/Project/pml-training.csv")[-1:-7]
testing = read.csv("education/practical machine learning/Project/pml-testing.csv")[-1:-7]
```

Number of training and testing samples and variables
```{r}
dim(training); dim(testing)
```
Remove columns where there is missing data (NA). Probably do not need to interpolate NA data since
have plenty of other variable with complete numeric data.

```{r}
retain <- colSums(is.na(training))==0
testing = testing[, retain]
training = training[, retain]
```
Now the dim is reduced
```{r}
dim(training);dim(testing)
```
Remove non numeric data from dataframe
```{r}
nums <- sapply(training, is.numeric)
```
But keep the manner in which they did the exercise aka the "classe" variable, 
which is the factor we are interested in predicting
```{r}
nums["classe"] = TRUE 
training <- training[, nums]
testing <- testing[, nums]
```
and the number of variable is considerably reduced
```{r}
dim(training); dim(testing)
```

Principal component analysis is used to find the most important features 
that explain 90 % of the data.
```{r}
pca <- preProcess(training[, -53], method="pca", thresh=0.90)
trainingPC <- predict(pca, training[, -53])
```
An example plot to see that the principal components are different for the first two results
```{r}
plot(trainingPC[,1],trainingPC[,2],col=training$classe)
```

Use the PC in a randomForest
```{r}
rf <- randomForest(training$classe ~ ., data=trainingPC)
testingPC <- predict(pca, testing[, -53])
```
Use the model on the test sample
```{r}
pre <- predict(rf, testingPC)
```
Print out the predictions for the test data set
```{r}
pre
```
Chart of variable importance as measured by Gini
```{r}
varImpPlot(rf, sort=TRUE, main="Variable importance for principal components")
```

Lets compare to using a randomForest using all the given variables
```{r}
rf <- randomForest(classe ~ ., data=training)


Use the model on the test samples
```{r}
pre <- predict(rf, testing)
pre
```
One point of data (3) in the testing set is different to the PCA data.
Chart the variable importance as measured by Gini
```{r}
varImpPlot(rf, sort=TRUE, main="Variable importance for training")
```









